---
title: "Report coursework assignment A - 2021"
subtitle: "CS4125 Seminar Research Methodology for Data Science"
author: "Nikki Bouman (4597648), Anuj Singh (), Gwennan Smitskamp ()"
date: "20/04/2021"
output:
   pdf_document:
      fig_caption: true
      number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


\tableofcontents

# Introduction
We will be using the following packages:
```{r}
library(ggplot2) # plotting
library(AICcmodavg) # aictab
library(pander) #for rendering output
library(rethinking) # for stan
```

# Part 1 - Design and set-up of true experiment 


## The motivation for the planned research
(Max 250 words)

## The theory underlying the research  
(Max 250 words) Preferable based on theories reported in literature


## Research questions 
The research question that will be examined in the experiment (or alternatively the hypothesis that will be tested in the experiment)


## The related conceptual model 
This model should include:
*Independent variable(s)
*Dependent variable
*Mediating variable (at least 1)
*Moderating variable (at least 1)


## Experimental Design 
Note that the study should have a true experimental design

## Experimental procedure 
Describe how the experiment will be executed step by step


## Measures
Describe the measure that will be used

## Participants
Describe which participants will recruit in the study and how they will be recruited

## Suggested statistical analyses
Describe the statistical test you suggest to care out on the collected data

<!-- # Part 2 - Generalized linear models -->

<!-- ## Question 1 Twitter sentiment analysis (Between groups - single factor)  -->

<!-- ### Conceptual model -->
<!-- Make a conceptual model for the following research question: Is there a difference in the sentiment of the tweets related to the different celebrities? -->

<!-- ### Collecting tweets, and data preparation -->
<!-- Include the annotated R script (excluding your personal Keys and Access Tokens information), but put echo=FALSE, so code is not included in the output pdf file. -->


<!-- ```{r, echo=FALSE, message=FALSE, warning=FALSE, include = FALSE} -->

<!-- #during writing you could add "eval = FALSE",  kntr will than not run this code chunk (take some time do) -->

<!-- #setwd("~/surfdrive/Teaching/own teaching/IN4125 - Seminar Research Methodology for Data Science/2019/coursework A")  -->
<!-- # commented this. Let's make this relative when trying to fix this file -->
<!-- # apple , note use / instead of \, which used by windows -->


<!-- #install.packages("twitteR", dependencies = TRUE) -->
<!-- library(twitteR) -->
<!-- #install.packages("RCurl", dependencies = T) -->
<!-- library(RCurl) -->
<!-- #install.packages("bitops", dependencies = T) -->
<!-- library(bitops) -->
<!-- #install.packages("plyr", dependencies = T) -->
<!-- library(plyr) -->
<!-- #install.packages('stringr', dependencies = T) -->
<!-- library(stringr) -->
<!-- #install.packages("NLP", dependencies = T) -->
<!-- library(NLP) -->
<!-- #install.packages("tm", dependencies = T) -->
<!-- library(tm) -->
<!-- #install.packages("wordcloud", dependencies=T) -->
<!-- #install.packages("RColorBrewer", dependencies=TRUE) -->
<!-- library(RColorBrewer) -->
<!-- library(wordcloud) -->
<!-- #install.packages("reshape", dependencies=T) -->
<!-- library(reshape) -->

<!-- ################### functions -->


<!-- clearTweets <- function(tweets, excl) { -->

<!--   tweets.text <- sapply(tweets, function(t)t$getText()) #get text out of tweets  -->


<!--   tweets.text = gsub('[[:cntrl:]]', '', tweets.text) -->
<!--   tweets.text = gsub('\\d+', '', tweets.text) -->
<!--   tweets.text <- str_replace_all(tweets.text,"[^[:graph:]]", " ") #remove graphic -->


<!--   corpus <- Corpus(VectorSource(tweets.text)) -->

<!--   corpus_clean <- tm_map(corpus, removePunctuation) -->
<!--   corpus_clean <- tm_map(corpus_clean, content_transformer(tolower)) -->
<!--   corpus_clean <- tm_map(corpus_clean, removeWords, stopwords("english")) -->
<!--   corpus_clean <- tm_map(corpus_clean, removeNumbers) -->
<!--   corpus_clean <- tm_map(corpus_clean, stripWhitespace) -->
<!--   corpus_clean <- tm_map(corpus_clean, removeWords, c(excl,"http","https","httpst")) -->


<!--   return(corpus_clean) -->
<!-- }  -->


<!-- ## capture all the output to a file. -->

<!-- ################# Collect from Twitter -->

<!-- # for creating a twitter app (apps.twitter.com) see youtube https://youtu.be/lT4Kosc_ers -->
<!-- #consumer_key <-'your key' -->
<!-- #consumer_scret <- 'your secret' -->
<!-- #access_token <- 'your access token' -->
<!-- #access_scret <- 'your access scret' -->

<!-- source("your_twitter.R") #this file will set my personal variables for my twitter app, adjust the name of this file. use the provide template your_twitter.R -->

<!-- setup_twitter_oauth(consumer_key,consumer_scret, access_token,access_scret) #connect to  twitter app -->


<!-- ##### This example uses the following 3 celebrities: Donald Trump, Hillary Clinton, and Bernie Sanders -->
<!-- ##  You should replace this with your own celebrities, at least 3, but more preferred  -->
<!-- ##  Note that it will take the computer some to collect the tweets -->

<!-- tweets_T <- searchTwitter("#trump", n=300, lang="en", resultType="recent") #300 recent tweets about Donald Trump, in English (I think that 1500 tweets is max) -->
<!-- tweets_C <- searchTwitter("#hillary", n=300, lang="en", resultType="recent") #300 recent tweets about Hillary Clinton -->
<!-- tweets_B <- searchTwitter("#bernie", n=300, lang="en", resultType="recent") #300 recent tweets about Bernie Sanders -->



<!-- ######################## WordCloud -->
<!-- ### This not requires in the assignment, but still fun to do  -->

<!-- # based on https://youtu.be/JoArGkOpeU0 -->

<!-- #corpus_T<-clearTweets(tweets_T, c("trump","amp","realdonaldtrump","trumptrain","donald","trumps","alwaystrump")) #remove also some campain slogans -->
<!-- #wordcloud(corpus_T, max.words=50) -->

<!-- #corpus_C<-clearTweets(tweets_C, c("hillary","amp","clinton","hillarys")) -->
<!-- #wordcloud(corpus_C,  max.words=50) -->

<!-- #corpus_B<-clearTweets(tweets_B, c("bernie", "amp", "sanders","bernies")) -->
<!-- #wordcloud(corpus_B,  max.words=50) -->
<!-- ############################## -->

<!-- ######################## Sentiment analysis -->

<!-- tweets_T.text <- laply(tweets_T, function(t)t$getText()) #get text out of tweets  -->
<!-- tweets_C.text <- laply(tweets_C, function(t)t$getText()) #get text out of tweets -->
<!-- tweets_B.text <- laply(tweets_B, function(t)t$getText()) #get text out of tweets -->



<!-- #taken from https://github.com/mjhea0/twitter-sentiment-analysis -->
<!-- pos <- scan('positive-words.txt', what = 'character', comment.char=';') #read the positive words -->
<!-- neg <- scan('negative-words.txt', what = 'character', comment.char=';') #read the negative words -->

<!-- source("sentiment3.R") #load algoritm -->
<!-- # see sentiment3.R form more information about sentiment analysis. It assigns a intereger score -->
<!-- # by substracitng the number of occurrence of negative words from that of positive words -->

<!-- analysis_T <- score.sentiment(tweets_T.text, pos, neg) -->
<!-- analysis_C <- score.sentiment(tweets_C.text, pos, neg) -->
<!-- analysis_B <- score.sentiment(tweets_B.text, pos, neg) -->


<!-- sem<-data.frame(analysis_T$score, analysis_C$score, analysis_B$score) -->


<!-- semFrame <-melt(sem, measured=c(analysis_T.score,analysis_C.score, analysis_B.score )) -->
<!-- names(semFrame) <- c("Candidate", "score") -->
<!-- semFrame$Candidate <-factor(semFrame$Candidate, labels=c("Donald Trump", "Hillary Clinton", "Bernie Sanders")) # change the labels for your celibrities -->

<!-- #The data you need for the analyses can be found in semFrame -->

<!-- ``` -->

<!-- ### Homogeneity of variance analysis -->
<!-- Analyze the homogeneity of variance of sentiments of the tweets of the different celebrities, and provide interpretation -->

<!-- ```{r} -->
<!-- #include your code and output in the document -->
<!-- ``` -->


<!-- ### Visual inspection Mean and distribution sentiments -->
<!-- Graphically examine the mean and distribution sentiments of tweets for each celebrity, and provide interpretation -->

<!-- ```{r} -->
<!-- #include your code and output in the document -->
<!-- ``` -->
<!-- ### Frequentist approach -->

<!-- #### Linear model -->
<!-- Use a linear model to analyze whether the knowledge to which celebrity a tweet relates has a significant impact on explaining the sentiments of the tweets. Provide interpretation of results  -->

<!-- ```{r} -->
<!-- #include your code and output in the document -->
<!-- ``` -->

<!-- #### Post Hoc analysis -->
<!-- If a model that includes the celebrity is better in explaining the sentiments of tweets than a model without such predictor, conduct a post-hoc analysis with e.g. Bonferroni correction, to examine which of celebrity tweets differ from the other celebrity tweets. Provide interpretation of the results -->

<!-- ```{r} -->
<!-- #include your code and output in the document -->
<!-- ``` -->

<!-- #### Report section for a scientific publication -->
<!-- Write a small section for a scientific publication, in which you report the results of the analyses, and explain the conclusions that can be drawn. -->

<!-- ### Bayesian Approach -->

<!-- #### Model description -->

<!-- Describe the mathematical model fitted on the most extensive model. (hint, look at the mark down file of the lectures to see example on formulate mathematical models in markdown). Justify the priors. -->

<!-- #### Model comparison -->

<!-- Conduct model analysis and provide brief interpretation of the results -->

<!-- ```{r} -->
<!-- #include your code and output in the document -->
<!-- ``` -->

<!-- #### Comparison celebrity pair -->

<!-- Compare sentiments of celebrity pairs and provide a brief interpretation (e.g. CIs)  -->


<!-- ## Question 2 - Website visits (between groups - Two factors) -->

<!-- ### Conceptual model -->
<!-- Make a conceptual model underlying this research question -->

<!-- ### Visual inspection -->
<!-- Graphically examine the variation in page visits for different factors levels (e.g. histogram, density plot etc.)  -->


<!-- ```{r} -->
<!-- #include your code and output in the document -->
<!-- ``` -->


<!-- ### Normality check -->
<!-- Visually inspect if variable page visits deviates from a Gaussian distribution, and discuss implication for general linear model analysis. -->


<!-- ```{r} -->
<!-- #include your code and output in the document -->
<!-- ``` -->

<!-- ### Frequentist Approach -->

<!-- #### Model analysis -->
<!-- Conduct a model analysis, to examine the added values of adding 2 factors and interaction between the factors in the model to predict page visits, and include brief interpretation of the results. -->


<!-- ```{r} -->
<!-- #include your code and output in the document -->
<!-- ``` -->


<!-- #### Simple effect analysis -->
<!-- If the analysis shows a significant two-way interaction effect, conduct a Simple Effect analysis to explore this interaction effect in more detail.It helps first to look at the means of different conditions in a figure. Provide brief interpretation of the results. -->


<!-- ```{r} -->
<!-- #include your code and output in the document -->
<!-- ``` -->


<!-- #### Report section for a scientific publication -->
<!-- Write a small section for a scientific publication, in which you report the results of the analyses, and explain the conclusions that can be drawn. -->

<!-- ### Bayesian Approach -->

<!-- #### Model description -->

<!-- Describe the mathematical model fitted on the most extensive model. (hint, look at the mark down file of the lectures to see example on formulate mathematical models in markdown). Justify the priors. -->

<!-- #### Model comparison -->

<!-- Conduct model analysis and provide brief interpretation of the results -->

<!-- ```{r} -->
<!-- #include your code and output in the document -->
<!-- ``` -->



# Part 3 - Multilevel model

## Visual inspection
<!-- Use graphics to inspect the distribution of the score, and relationship between session and score -->
The boxplot and density plot show the distribution of the score. We can see that the mean score is 122 points. The minimum is set at 59, with outliers until 46, while the maximum is set at 190, with outliers until 248.

```{r}
# Get data
filepath <- ("set0.csv")
ds <- read.csv(file=filepath, header=TRUE)
ds <- data.frame(ds)

# boxplot score overall distribution (session independent)
boxplot(ds$score)

# density score overall distribution (with mean line)
p <- ggplot(ds, aes(x=score)) + geom_density()
p + geom_vline(aes(xintercept=mean(score)), color="blue", linetype="dashed",size=1)
```

The relationship between the score and the session can be observed with the next two figures. The regression line (blue) in the scatterplot clearly shows how the score rises with the amount of sessions. This can also be observed in the box plot when looking at the mean (black line) for every box.

```{r}
# set labels
ds$sessionF <- factor(ds$session, levels=c(0:49), labels=c(0:49))

# boxplot score per session
boxplot(score~sessionF, data=ds, main="Score", xlab="session", ylab="score")

# ggplot score per session
hp <- ggplot(ds, aes(x=session, y=score)) + geom_point(shape=1) + geom_smooth(formula = y ~ x,method=lm)
hp
```

## Frequentist approach

### Multilevel analysis
<!-- Conduct multilevel analysis and calculate 95% confidence intervals, determine: -->
We have conducted a multilevel analysis. We have an intercept only model (model0) which we compare to a model that includes a predictor parameter for the session (model1). By comparing these two models, we will know whether there is a difference in the score over the sessions.

```{r}
# create models as given in slides lecture 4
model0 <- lm(formula=score~1, data=ds, na.action=na.exclude)
model1 <- lm(formula=score~sessionF, data=ds, na.action=na.exclude)

# analysis, see if predictor improves fitting
pander(anova(model0,model1))
pander(anova(model1))
```

From this analysis we can see there is a significant variation between the sessions. We take a further look at the summary results.

```{r}
pander(summary(model1))
```

The summary compares the first session (intercept) with the other sessions. Looking at the estimates, we can see that compared to the first session, the scores are higher every later session. Next, we will take a look at the Akaike Information Criterion (AIC) to compare the models on the goodness-of-fit concering the out-of-sample deviance.

```{r}
#AIC
models <- list(model0, model1)
model.names <- c("model0", "model1")
pander(aictab(cand.set = models, modnames=model.names), caption="Model selection based on AICc.")
```

Here we can see that model 1 has the best goodness-of-fit as it has the smallest AICc value. Lastly, we will obtain a 95% confidence interval of the estimates we have obtained earlier.

```{r}
# gives CI95%
pander(confint(model1), caption="95% confidence interval of the estimates.")
```

<!-- * If session has an impact on people score -->
<!-- * If there is significant variance between the participants in their score -->
Here we can again see an increase in score related to the sessions. From this we can conclude that the session has a positive effect on people's score. Also, it seems there is a significant variance between the participants in their score.

### Report section for a scientific publication
<!-- Write a small section for a scientific publication, in which you report the results of the analyses, and explain the conclusions that can be drawn. -->

A Linear Model analysis was conducted to test the difference between sessions on the score. The results found a significant effect (F(49,16078) = 52.793, p < .001) for the sessions on the score. We can conclude that over sessions the score per participant generally increases.

## Bayesian approach

### Model description
<!-- Describe the mathematical model fitted on the most extensive model. (hint, look at the mark down file of the lectures to see example on formulate mathematical models in markdown). Justify the priors. -->

For model 2, the model with session as a factor, we take as prior a normal distribution of N(125,30). This comes from the mean of the score, 125, and a bit more than the standard deviation, which is around 27. Our sigma is set at a uniform distribution of U(0.001,30).

### Model comparison
<!-- Select the first 100 participants from the data set. (hint to overcome the Stan problem with a zero index, increase subject id number with 1). Compare models with with increasing complexity.  -->

We will create and compare the three described models. From the results we can see that m1 has the best fit since it has the smallest WAIC value

```{r}
ds <- ds[!(ds$Subject>99),] # select first 100 subjects
ds$Subject <- ds$Subject +1 # increase subject number with 1 to overcome Stan zero index problem

mean(ds$score) # check mean
sd(ds$score) # check standard deviation

ds$sessionF <- factor(ds$session, levels=c(0:49), labels=c(0:49))
ds$subjF <- factor(ds$Subject, levels=c(1:100), labels=c(1:100))

da <- subset(ds, select=c(score, sessionF))
da1 <- subset(ds, select=c(score, subjF))

# create model with fixed intercept (i)
m0 <- map2stan(
  alist(
    score ~ dnorm(mu, sigma),
    mu <- a,
    a ~dnorm(125,30), # mean and sd from what we found above
    sigma ~dunif(0.001,30)
  ), data = da, iter = 10000, chains = 4, cores = 4
)

# create model extended with an adaptive prior for subject id (ii)
m1 <- map2stan(
  alist(
    score ~ dnorm(mu, sigma),
    mu <- a[subjF],
    a[subjF] ~ dnorm(125,sigma_subj),
    sigma_subj ~ dcauchy(0,10),
    sigma ~dunif(0.001,30)
  ), data = da1, iter = 10000, chains = 4, cores = 4
)

# create model with session as a factor (iii)
m2 <- map2stan(
  alist(
    score ~ dnorm(mu, sigma),
    mu <- a[sessionF],
    a[sessionF] ~ dnorm(125,30),
    sigma ~dunif(0.001,30)
  ), data = da, iter = 10000, chains = 4, cores = 4
)

pander(compare(m0,m1,m2,func=WAIC))
```

### Estimates examination
<!-- Examine the estimate of parameters of the model with best fit, and provide a brief interpretation. -->
From the previous comparison we could see that m1 is the best fit model. We will further examine this model with 95% credible intervals of the parameters of this model.

```{r}
pander(precis(m1, depth=2, prob=.95))
```

Looking at the mean we can see that in general the scores are increasing over the sessions. Moreover, the standard deviating is increasing as well. This could be due to the fact that not every subject has completed the same amount of sessions.


